### 2023-03-22 S1_replicate_stats_and_plots.Rmd
basic analyses ready

---
title: 'Suplementary Material 1: Code to Reproduce Analysis and Plots'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 2)
options(scipen = 9)
options(knitr.kable.NA="")
```

## Introduction

The aim of this document is to display the R code needed the reproduce the findings of the main text. This documents needs the twinR package to be installed to run.

```{r}
#cleanup memory
gcstuff <- gc(verbose=FALSE); rm(gcstuff);
```

```{r}
#get last birth adding function
source("./R/last_birth.R")

#simplified twinR summary tables 
source("./R/twinR_summary.R")

#fix twinR compute predictions to do prediction with no lambda  as well
source("./R/twinR_predictions.R")

#simple convenience functions
source("./R/utils.R")
```

```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- min(c(50L, parallel::detectCores() - 1))

## Set option in spaMM:
spaMM::spaMM.options(nb_cores = nb_cores)
```

## Data import

The Estonian dataset has been formatted to include the same columns as *the data_births_all* dataset from the **twinR** package. The only difference is that the columns *pop* and *monthly* are excluded as these are constant.

```{r}
#Import and preproccess Estonian Data

data_births_monthly_EE <- readRDS("./data/data_births_all_EE.rds")

#the twinR package expects population to be present
data_births_monthly_EE$pop <- "Estonia"

## Expand the birth level data for the fit of statistical models:
data_births_monthly_EE <- twinR::expand_data(data_births_monthly_EE)

data_births_monthly_EE <- add_last_birth(data_births_monthly_EE)

data_births_monthly_EE_not_last <- data_births_monthly_EE[!data_births_monthly_EE$last,]

#make the aggregates
dmm_EE <- twinR::aggregate_data(data_births_monthly_EE)
dmm_EE$prob_twin <- dmm_EE$twin_total / dmm_EE$births_total

dmm_EE_nl <- twinR::aggregate_data(data_births_monthly_EE_not_last)
dmm_EE_nl$prob_twin <- dmm_EE_nl$twin_total / dmm_EE_nl$births_total
```

```{r}
ee_tbl <- twinR::build_data_summary.table(data_births_monthly_EE)
ee_tbl_nl <- twinR::build_data_summary.table(data_births_monthly_EE_not_last)
```

\tiny
```{r}

  eetbls <- rbind(ee_tbl[-1,4:ncol(ee_tbl)-1],ee_tbl_nl[-1,4:ncol(ee_tbl_nl)-1])
eetbls$Subset <- c("All data", "No last birth")
knitr::kable(eetbls,
             caption = paste0("Data summary table of Estonian data",
                              " with and without last born child")) 
```
\normalsize

## Data import Original

```{r}
##Import and pre-proccess twinR package data

## Filter the raw data to only keep data with monthly resolution:
data_births_monthly <- twinR::filter_data(twinR::data_births_all)

## Expand the birth level data for the fit of statistical models:
data_births_monthly <- twinR::expand_data(data_births_monthly)

data_births_monthly <- add_last_birth(data_births_monthly)
data_births_monthly_not_last <- data_births_monthly[!data_births_monthly$last,]

dmm_orig <- twinR::aggregate_data(data_births_monthly)
dmm_orig$prob_twin <- dmm_orig$twin_total /dmm_orig$births_total

dmm_orig_nl <- twinR::aggregate_data(data_births_monthly_not_last)
dmm_orig_nl$prob_twin <- dmm_orig_nl$twin_total /dmm_orig_nl$births_total

```

\tiny
```{r}
knitr::kable(twinR::build_data_summary.table(data_births_monthly_not_last)[-2],
             caption = "Data sumary table without last born child")
```
\normalsize

## Fitting models

```{r}
#' Fit Predictions
#'
#' This function fits a model using the given formula and dataset and 
#' computes predictions. The model is fit using the \code{\link[spaMM]{fitme}}
#' function from the \strong{spaMM} package.
#' 
#' @param dataset A data frame containing the data to be used for fitting the model.
#' @param formula A formula specifying the model to be fit.
#' @param predict Logical value indicating whether to do predictions. Default is TRUE.
#' @param nb_boot Number of bootstrap samples to use when computing predictions.
#'  Default is 1000.
#' @param predictionsDir Directory where precomputed predictions are stored.
#'  Default is "./data/predictions".
#' 
#' @return A list containing the fitted model object and a data frame with
#'  computed predictions.
fitPredictions <- function(dataset, formula,
                           predict=TRUE,
                           nb_boot=1000,
                           predictionsDir = "./data/predictions"){

  if(!dir.exists(predictionsDir)) dir.create(predictionsDir)

  args <- list(formula = stats::as.formula(formula),
             data = dataset, 
             family = stats::binomial(link = "logit"),
             method = "PQL/L")

  fit <-  twinR::fit_model_safely(timeout = Inf, .args = args)

  fitName <- deparse(substitute(dataset))
  #TODO maybe add formula also to the fitName like:
  #fitName <- paste0(form2str(stats::as.formula(formula)),fitName)

  predDataFname <- paste0(predictionsDir,"/",fitName ,"data_fig.rds")
  if(!file.exists(predDataFname) & predict){
    min_births <- min(dataset$births_total)
    max_births <- max(dataset$births_total)
    nd <- data.frame(births_total = min_births:max_births)
    data_fig <- compute_predictions(fit,
                                    newdata = nd,
                                    nb_boot = nb_boot)

    saveRDS(data_fig, predDataFname)

  } else {
    if(file.exists(predDataFname)){
      warning("Pre-computed predictions returned from file:\n", predDataFname,
            "\n If you want to re-run this time intensive step delete the file!")
    data_fig <- readRDS(predDataFname)
    } else {
      data_fig <- list(results = NULL)
    }

  }

  #garbage collecion after spaMM multi-core proccess
  gcstuff <- gc(verbose=FALSE)

  return(list(fit=fit, results=data_fig$results))
}
```

### Full Data

```{r}
## Estonia - mother level data
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_fit <- fitPredictions(dmm_EE, formula)

knitr::kable(build_fit_summary.table(dmm_EE_fit$fit))
```

```{r}
## Estonia - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id)"
dbm_EE_fit <- fitPredictions(data_births_monthly_EE, formula, predict = F)

knitr::kable(build_fit_summary.table(dbm_EE_fit$fit))
```

```{r}
## TwinR - mother level data
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + (1|pop)"
dmm_orig_fit <- fitPredictions(dmm_orig, formula)

knitr::kable(build_fit_summary.table(dmm_orig_fit$fit))
```

```{r}
## TwinR - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id) + (1|pop)"
dbm_orig_fit <- fitPredictions(data_births_monthly, formula, predict = F)

knitr::kable(build_fit_summary.table(dbm_orig_fit$fit))
```

### No Last Births Data

```{r}
## Estonia
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total"
dmm_EE_nl_fit <- fitPredictions(dmm_EE_nl, formula)

knitr::kable(build_fit_summary.table(dmm_EE_nl_fit$fit))
```

```{r}
## Estonia - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id)"
dbm_EE_nl_fit <- fitPredictions(data_births_monthly_EE_not_last, formula,
                                predict = F)

knitr::kable(build_fit_summary.table(dbm_EE_nl_fit$fit))
```

```{r}
## TwinR
formula <- "cbind(twin_total, singleton_total) ~ 1 + births_total + (1|pop)"
dmm_orig_nl_fit <- fitPredictions(dmm_orig_nl, formula)

knitr::kable(build_fit_summary.table(dmm_orig_nl_fit$fit))
```

```{r}
## TwinR - birth level data
formula <- "twin ~ 1 + poly(cbind(age, parity), 3) + (1|maternal_id) + (1|pop)"
dbm_orig_nl_fit <- fitPredictions(data_births_monthly_not_last, formula, predict = F)

knitr::kable(build_fit_summary.table(dbm_orig_nl_fit$fit))
```

## Plots

```{r}
library(ggplot2)
#some nice colors
bc <- c("purple", "black", "navy", "darkgoldenrod2", "springgreen3", "gray")

#use new base theme that displays also grid lines 
source("./R/twinR_theme.R")
```

### Fig 1: Estonian vs TwinR Full Data

```{r}
fig2_EE_plot_data <- dmm_EE_fit$results
fig2_orig_plot_data <- dmm_orig_fit$results
```

```{r}

fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color="EE all"),  size = 1) +
    stat_summary(data=dmm_EE[dmm_EE$births_total <17, ],
                 aes(x=births_total, y=prob_twin, color="EE all", fill = "EE all"),
                 alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color="EE all", fill = "EE all"),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data,
             aes(y = estimate,  x=births_total, color="orig. rural"),  size = 1) +
    stat_summary(data=dmm_orig[dmm_orig$births_total <19, ],
                 aes(x=births_total, y=prob_twin,
                     color="orig. rural", fill="orig. rural"),
                 alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_orig_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill="orig. rural"),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Model prediction + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "dataset")

p2 
```

### Fig 2: Estonian vs Others No Last Birth

```{r}
fig2_EE_plot_data_nl <- dmm_EE_nl_fit$results
fig2_EE_plot_data_nl$births_total <- fig2_EE_plot_data_nl$births_total + 1
fig2_orig_plot_data_nl <- dmm_orig_nl_fit$results
fig2_orig_plot_data_nl$births_total <- fig2_orig_plot_data_nl$births_total + 1
```

```{r}
dmm_EE_nl_plot <- dmm_EE_nl
dmm_orig_nl_plot <- dmm_orig_nl
dmm_EE_nl_plot$births_total <- dmm_EE_nl_plot$births_total + 1
dmm_orig_nl_plot$births_total <- dmm_orig_nl_plot$births_total + 1
```

```{r}
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data_nl,
              aes(y = estimate,  x=births_total, color="EE all"),  size = 1) +
    stat_summary(data=dmm_EE_nl_plot,
                 aes(x=births_total, y=prob_twin, color="EE all", fill = "EE all"),
                 alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data_nl,
                aes(y = estimate, x=births_total, ymin = lwr, ymax = upr,
                    color="EE all", fill = "EE all"),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data_nl,
             aes(y = estimate,  x=births_total, color="orig. rural"),  size = 1) +
    stat_summary(data=dmm_orig_nl_plot,
                 aes(x=births_total, y=prob_twin,
                     color="orig. rural", fill="orig. rural"),
                 alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_orig_plot_data_nl,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill="orig. rural"),
                alpha = 0.3) +
    ggplot2::scale_x_continuous(breaks = 1:18, limits = c(1,NA)) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.005)) +
    ggplot2::coord_cartesian(ylim=c(0,0.03)) +
    labs(subtitle = "Without last children, model prediction + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p3 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "dataset")

p3 
```

```{r}
#END
```