### 2023-03-22 test.Rmd

---
title: "Test"
author: "Richard Meitern"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is an R Markdown test document.

```{r cars}
knitr::kable(summary(cars))
```

## Including Plots
```{r}
library(ggplot2)
```

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ggplot(pressure, aes(x=temperature, y=pressure)) + geom_point()
```

```{r}
#END
```