### 2023-04-04 test.Rmd

---
title: "Test"
author: "Richard Meitern"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is an R Markdown test document.

```{r cars}
testDir <- "./exports/test/"
knitr::kable(summary(cars))
dir.create(testDir, recursive = T)
writexl::write_xlsx(cars, "./exports/test/test_cars.xlsx")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
ggplot2::ggplot(pressure, ggplot2::aes(x=temperature, y=pressure)) + ggplot2::geom_point()
```

```{r}
#END
```