### 2023-04-18 S4_draw_IBI_predictions_EE_not_last.Rmd
replicate fig 4

---
title: 'Suplementary Material 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data'
author: "Richard Meitern"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!is.null(knitr::opts_knit$get("rmarkdown.pandoc.to"))){
 if (knitr::opts_knit$get("rmarkdown.pandoc.to") == "docx"){
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
} 
}

options(digits = 9)
options(scipen = 9)
options(knitr.kable.NA="")
```

## Introduction

The aim of this document is to display the R code needed the reproduce the original paper figure 4 equivalent on the Estonian dataset where the last births have been ommited.

**NB!** Some of the code e.g reading in the data is same as in S1.

```{r}
#cleanup memory
gcstuff <- gc(verbose=FALSE); rm(gcstuff);

```

```{r}
cat("The current directory is: ", getwd())
```

```{r}
#get last birth adding function
source("./R/last_birth.R")

#simplified twinR summary tables 
source("./R/twinR_summary.R")

#fix twinR compute predictions to do prediction with no lambda  as well
source("./R/twinR_predictions.R")

#simple convenience functions
source("./R/utils.R")
```

```{r}
## Identify number of CPU cores available for parallel computing,
## note: using a large number may lead RAM to max out, so you may have to adjust
## that according to your infrastructure:
nb_cores <- 7

#it seems that simulating the base model for Estonian not last birth data
#requires about 15GB of RAM per CPU core (tested with 61 CPU cores)

## Set option in spaMM:
spaMM::spaMM.options(nb_cores = nb_cores)
```

## Data Import Estonia

The Estonian dataset has been formatted to include the same columns as *the data_births_all* dataset from the **twinR** package. The only difference is that the columns *pop* and *monthly* are excluded as these are constant.

```{r}
#Import and preproccess Estonian Data

data_births_monthly_EE <- readRDS("./data/data_births_all_EE.rds")

#the twinR package expects population to be present
data_births_monthly_EE$pop <- "Estonia"

data_births_monthly_EE <- add_last_birth(data_births_monthly_EE)

data_births_monthly_EE_not_last <- data_births_monthly_EE[!data_births_monthly_EE$last,]

```

## Get Models

```{r}
fig_data_files <- sapply(LETTERS[1:3], function(l){
  predDir <- "./data/predictions/"
  paste0(predDir, "data_fig_4",l, ".rds")
})
```

```{r}
if(!all(sapply(fig_data_files, file.exists))){
  fitDir <- "D:/hpc/projects/natCom_twinR_commentary/exports/fitted_models_EE_nl/"
  load(paste0(fitDir, "fits_PISH_obs.rda")) # fit all

  data_fig_4A <- twinR::prepare_data_fig_4A(fits$fit_PP)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4A, paste0(predDir, "data_fig_4A", ".rds"))

  data_fig_4B <- twinR::prepare_data_fig_4B(fits$fit_IBI)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4B, paste0(predDir, "data_fig_4B", ".rds"))

  data_fig_4C <- twinR::prepare_data_fig_4C(fits$fit_twinning.binary)
  gcs <- gc(verbose = F)
  saveRDS(data_fig_4C, paste0(predDir, "data_fig_4C", ".rds"))
} else{
  for(l in LETTERS[1:3]){
    assign(paste0("data_fig_4",l), readRDS(fig_data_files[l]))
    }
}

```

## Plot models

```{r}
fig_4A <- twinR::draw_fig_4A(data_fig_4A)
fig_4B <- twinR::draw_fig_4B(data_fig_4B)
fig_4C <- twinR::draw_fig_4C(data_fig_4C)
```

```{r}
twinR::plot_grid(fig_4A, fig_4B, fig_4C,
          labels = "auto", nrow = 1, label_size = 7, align = "v", axis = "lr")
```

```{r}
ggplot2::ggsave(file = "./exports/fig4.pdf", width = 180, height = 55, units = "mm")
```

```{r}
#END
```